---
title: "Customer Analysis2"
author: "*Amazing*"
date: "*2/6/2020*"
output:
  html_notebook:
    theme: journal
    highlight: zenburn
editor_options:
  chunk_output_type: inline
---

</br></br>
*<font size="5">Step1: Import the data </font>*
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(BTYD)
# load calib data
calib <- read.table("cruises_calib.txt",header = T)
# load holdout data
hold <- read.table("cruises_hold.txt",header = T)

```
</br></br>
*<font size="5">Step2: Clean data </font>*
```{r}
calib$Year <- calib$Year-2009
recency <- calib %>% group_by(ID) %>% summarise(recency = max(Year))
calib <- cbind(calib, value = rep(1,nrow(calib)))
calib <- calib %>%
  spread(key="Year",value="value")
calib[is.na(calib)] <- 0
calib <- merge(cbind(calib, rowSums(calib[,3:7])),recency, by = "ID")
colnames(calib) <- c("customerID","2009","2010","2011","2012","2013","2014","frequency","recency")
NumTot <- nrow(calib)
calib_combine <- calib %>% count(frequency,recency)
```

</br></br>
*<font size="5">Question1: Build BB and BG/BB model and estimate parameters </font>*
```{r}
log_bb <- function(pars, x, m) {
  a <- pars[1]
  b <- pars[2]
  lchoose(m, x) + lbeta(a+x, b+m-x) - lbeta(a,b)
}

LL_BB <-function(pars, x, m, n) {
  p <- exp(pars)
  LL_ind <- log_bb(p, x, m)
  return(-sum(LL_ind * n))
}
```
</br></br>
*<font size="4">Test LL_BB from different starting point </font>*
```{r}
# test LL_BB from different starting point 
test <- tibble(a_start=c(0,1,2,10),b_start=c(0,1,2,10))
#return LL value
opt_value <- function(par){
  opt_BB <- optim(par, fn=LL_BB, x=calib_combine$frequency,m=5, n=calib_combine$n)
  return(-opt_BB$value)
}

#return a
opt_a <- function(par){
  opt_BB <- optim(par, fn=LL_BB, x=calib_combine$frequency,m=5, n=calib_combine$n)
  return(exp(opt_BB[["par"]][1]))
}

#return b
opt_b <- function(par){
  opt_BB <- optim(par, fn=LL_BB, x=calib_combine$frequency,m=5, n=calib_combine$n)
  return(exp(opt_BB[["par"]][2]))
}

test <- test %>% mutate(a=apply(test,1,FUN=function(x) opt_a(x)),
              b=apply(test,1,FUN=function(x) opt_b(x)),
              LL_BB=apply(test,1,FUN=function(x) opt_value(x)))
test
```
*<font size="4"> LL_BB are the same when we test data from different starting point, thus we choose a = 0, b = 0 as our starting point. </font>*

</br></br>
*<font size="5"> BB Model and parameter estimates with a = 0 and b = 0 as starting point </font>*
```{r}
pairs <- c(0,0)

opt_BB <- optim(pairs, fn=LL_BB, x = calib_combine$frequency, m = 5, n=calib_combine$n)
params_bb <- exp(opt_BB[["par"]])
LL_bb <- -opt_BB[["value"]]
params_bb
LL_bb
```
</br>
*<font size="4"> The parameters of BB model are a = 0.903133 and b = 9.0526529. They are the estimated parameters of beta distribution which describes heterogeneity in the probability of repeating cruise in the following years.Since a<1,b>1,so mixing distribution of p has a ‘L’ shape, suggesting that people are more likely to choose to repeat with frequency less than 5 times. </font>*

</br></br>
*<font size="5"> BG/BB Model </font>*
```{r}
LL_BGBB <- function(pars, x, tx, m, n) {
  p <- exp(pars)
  k <- pmax(0,lchoose(tx-1,x-1))
  LL_ind <- bgbb.LL(p, x, tx, m) + k
  return(-sum(LL_ind * n))
}
pairs <- c(0,0,0,0)
opt_BGBB <- optim(pairs, fn=LL_BGBB, x=calib_combine$frequency, tx=calib_combine$recency, m=5, n=calib_combine$n, control = list(maxit = 100000))
params_bgbb <- exp(opt_BGBB[["par"]])
LL_bgbb <- -opt_BGBB[["value"]]
params_bgbb
LL_bgbb
```
</br>
*<font size="4">The parameters of BG-BB model is a = 1.6603784,b = 11.6294724,c = 0.2441056 and d = 1.3174915. a and b are the parameters of beta distribution which describes heterogeneity in the probability of repeating cruise in the following years.c and d are the parameters of beta distribution which describes heterogeneity in the “death rates” of different customers. As a > 1, b > 1, beta distribution is mound-shaped,suggesting people are more likely has similar probalility of repeating cruise. As c < 1 and d > 1, beta distribution is L-shaped </font>*

</br></br> 
*<font size="5">Question2: Create both table and bar plot to demonstrate comparison.</font>*
```{r}
# observation vs. prediction
nbr_calib <- calib %>% count(frequency)
colnames(nbr_calib) <- c("x", "N")
nbr_calib <- cbind(nbr_calib, Pred_N_BB = round(exp(log_bb(params_bb,0:5,5))*NumTot,0), 
             Pred_N_BGBB = round(bgbb.pmf(params_bgbb,5,0:5)*NumTot,0))
nbr_calib


select(nbr_calib,x,Obs=N,Pred_N_BB, Pred_N_BGBB) %>%
  gather(Model,Count,c(Obs,Pred_N_BB, Pred_N_BGBB)) %>%
  ggplot(aes(x=x,y=Count,fill=Model,color=Model)) %>%
  + geom_bar(position="dodge",stat="identity") %>%
  + ggtitle("Calibration Data: Obs vs. Pred") %>%
  + scale_fill_manual(values=c('red','green','darkblue')) %>%
  + scale_color_manual(values=c('red','green','darkblue')) %>%
  + scale_y_continuous("Respondents") %>%
  + theme_bw() %>%
  + theme(plot.title = element_text(hjust = 0.5),
          text=element_text(size=15),
          legend.position="bottom")
```


```{r}
# discrepency comparison 
dis_calib <- nbr_calib %>%
  mutate(Dis_BB=(nbr_calib$N-nbr_calib$Pred_N_BB)/nbr_calib$N,Dis_BGBB=(nbr_calib$N-nbr_calib$Pred_N_BGBB)/nbr_calib$N)

select(dis_calib,x,Dis_BB,Dis_BGBB) %>%
  gather(Model,Count,c(Dis_BB,Dis_BGBB)) %>%
  ggplot(aes(x=x,y=Count,fill=Model,color=Model)) %>%
  + geom_line(stat="identity") %>%
  + ggtitle("Discrepancy Percentage") %>%
  + scale_fill_manual(values=c('green','darkblue')) %>%
  + scale_color_manual(values=c('green','darkblue')) %>%
  + scale_y_continuous("Respondents") %>%
  + theme_bw() %>%
  + theme(plot.title = element_text(hjust = 0.5),
          text=element_text(size=15),
          legend.position="bottom")
```

*<font size="4"> From the bar chart and discrepancy percentage graph, BB and BGBB predicts similar results during 2009 to 2014, because both models are trying to maximize fitness with existing data.</font>*

</br></br> 
*<font size="5">Question3: Discuss two models when predicting hold-out periods.</font>*

```{r}
# data cleaning and manipulation
hold$Year <- hold$Year-2014
recency <- hold %>% group_by(ID) %>% summarise(recency = max(Year))
hold <- cbind(hold, value = rep(1,nrow(hold)))
hold <- hold %>%
  spread(key="Year",value="value")
id <- data.frame(ID = seq(length(calib$customerID)))
hold <- merge(id, hold, by = "ID", all.x = T)
hold[is.na(hold)] <- 0
hold <- merge(cbind(hold, rowSums(hold[,2:5])),recency, by = "ID", all.x = T)
hold[is.na(hold)] <- 0
colnames(hold) <- c("customerID","2015","2016","2017","2018","frequency","recency")
hold_agg <- hold %>% count(frequency,recency)
```


```{r}
# observation vs. prediction
nbr_hold <- hold %>% count(frequency)
colnames(nbr_hold) <- c("x", "N")
nbr_hold <- cbind(nbr_hold, Pred_N_BB = round(exp(log_bb(params_bb,0:4,4))*NumTot,0), 
                  Pred_N_BGBB = round(bgbb.pmf.General(params_bgbb,5,4,0:4)*NumTot,0))
nbr_hold

select(nbr_hold,x,Obs=N,Pred_N_BB, Pred_N_BGBB) %>%
  gather(Model,Count,c(Obs,Pred_N_BB, Pred_N_BGBB)) %>%
  ggplot(aes(x=x,y=Count,fill=Model,color=Model)) %>%
  + geom_bar(position="dodge",stat="identity") %>%
  + ggtitle("Holdout Data: Obs vs. Pred") %>%
  + scale_fill_manual(values=c('red','green','darkblue')) %>%
  + scale_color_manual(values=c('red','green','darkblue')) %>%
  + scale_y_continuous("Respondents") %>%
  + theme_bw() %>%
  + theme(plot.title = element_text(hjust = 0.5),
          text=element_text(size=15),
          legend.position="bottom")
```
*<font size="4"> For holdout period, BGBB model predicts better than BB model from 2015 to 2018. As BGBB model considers both customer choosing behavior and death probability, while BB model only considers customer choosing behavior.</font>*

</br></br> 
*<font size="5">Question4: (a) Number of customer aquired at 2009 who are still alive at start of 2019. (b) Total expected residual lifetime value of 2009 custemers at 2019. </font>*
```{r}
# merge calib and hold
df <- merge(calib[,1:7], hold[,1:5], by = "customerID")
df <- cbind(df, rowSums(df[,3:11]))
df <- df %>% mutate(df_recn = apply(df[2:11], 1, function(x) {
                   max(x * seq(0,9))
                 }))
#df <- cbind(df, df_recn)
colnames(df)[c(12,13)] <- c("frequency","recency")
df_agg <- df %>% count(frequency,recency)

# expected number of alive customers (Question.a)
alive_exp <- round(sum(bgbb.PAlive(params_bgbb, df_agg$frequency, df_agg$recency, 9) * df_agg$n),0)
alive_exp
```


```{r}
# total expected residual transactions
delta <- 0.13
DERT_exp <- sum(bgbb.DERT(params_bgbb, df_agg$frequency, df_agg$recency, 9, delta) * df_agg$n)

# total expected residual lifetime value (Question.b)
ERLV_exp <- DERT_exp * 241
ERLV_exp
```

</br></br> 
*<font size="5">Question5: (a) Predicting 1193 new customer from 2020 to 2024. (b) Expected life time value of new customers </font>*
```{r}
#  Question.a: incremental transactions by year
new_cust <- 11993
incre_trans <- round(diff(bgbb.Expectation(params_bgbb, 0:5) * new_cust),0)
year_cnt <- data.frame(year = c("2020","2021","2022","2023","2024"), incremental = incre_trans)
year_cnt
```


```{r}
ggplot(year_cnt,aes(x = year,y = incremental,group = 1)) %>%
  + geom_line(color = "blue",size = 1) %>%
  + geom_point(color = "darkblue",size = 4) %>%
  + ggtitle("Incremental Transactions by Year") %>%
  + theme_bw() %>%
  + expand_limits(y = 700) %>%
  + geom_text(aes(label = incremental),hjust = 0.5,vjust = 2) %>%
  + theme(plot.title = element_text(hjust = 0.5),
          text=element_text(size = 15),
          legend.position = "bottom")
```


```{r}
# Question.b: the maximum spend on the campaigns, in other words, the ECLV
max_camp <- (bgbb.DERT(params_bgbb, 0, 0, 0, delta) + 1) * new_cust*241
max_camp
```
